---
title: "Homework 2"
author: "Shangsi Lin"
date: "2022/10/01"
output: github_document

---

```{r load_libraries, message = FALSE}
library(tidyverse)
library(readxl)
```

## Problem 1

Import data from `NYC_Transit_Subway_Entrance_And_Exit_Data.csv`. Then clean variable names and selects the interested column of variables. Using the `recode()` argument to change the entry variable from `YES` / `NO` to a logical variable. Also change the route 8-11 data type from double to character as the route 1-7 data.

```{r chunk_clean}
transit_df = 
  read_csv(
    "data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) %>% 
  mutate(entry = recode(entry, "YES" = TRUE, "NO" = FALSE))
```

These data are not tidy, after the cleaning above, the current dataset contains variables including line, station name, station latitude, station longitude, route information, entry, exit only or not, vending, entrance type, and ADA compliance. The dataset dimension is 1868 x 20. It would be helpful to change rount numbers and route as variables for investigations interested on specific routes.

To find out the number of unique stations, select station name and line and use `distinct()` to find all unique combinations that each represents a unique station. The number of rows represent the number of unique stations.

```{r chunk_unique}
transit_df %>% 
  select(station_name, line) %>% 
  distinct
```

Turns out, there are 465 unique stations.

To find out the number of ADA compliant stations, filter `ada == TRUE` first and then select unique stations the same as before. The number of rows represent the number of ADA compliant stations.

```{r chunk_ADA}
transit_df %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

Turns out, there are 84 ADA compliant stations.

To find out the proportion of station entrances / exits without vending allow entrance, first filter stations that don't allow vending, and then take the mean of the `entry` variable. The result is the parameter of interest.

```{r chunk_proportion}
transit_df %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

Turns out, the proportion of station entrances / exits without vending allow entrance is about 0.377

Reformat data so that route number and route name are distinct variables, then identify stations that serve the A train and how many of those are ADA compliant using the `filter`, `select`, and `distinct` arguments.

```{r chunk_reformat}
reformat_df = transit_df %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_number",
    values_to = "route") 

reformat_df %>%   
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

reformat_df %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

Turns out, there are 60 stations that serve the A train and 17 stations that serve the A train while being ADA compliant.

## Problem 2

Import two data sheets from `Trash-Wheel-Collection-Totals-7-2020-2.xlsx`, clean the names od the variables and filter out only dumpster-specific data. Round the number of sports balls to the nearest integer and converts the result to an integer variable. Rename variable name "dumpster" into "dumpster_number" for clarity. Add a new variable called `wheel_name` to keep track of which trash wheel does the data belong to.

```{r chunk_clean2}
mr_trash_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = 1, range = "A2:N534") %>% 
  janitor::clean_names() %>% 
  filter(dumpster != "NA") %>% 
  mutate(sports_balls = as.integer(round(sports_balls))) %>% 
  mutate(wheel_name = "mr_trash") %>% 
  rename(dumpster_number = dumpster)

professor_trash_df = 
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = 2, range = "A2:N116") %>%
  janitor::clean_names() %>% 
  filter(dumpster != "NA") %>% 
  mutate(sports_balls = as.integer(round(sports_balls))) %>% 
  mutate(wheel_name = "professor_trash") %>% 
  rename(dumpster_number = dumpster)
```

Combine both data sheets to produce a single tidy dataset, .

```{r chunk_join}
trash_wheel_df = 
  full_join(mr_trash_df, professor_trash_df, by = "wheel_name")
```

```{r chunk_calculation2}
total_weight = pull(professor_trash_df, weight_tons) %>% 
  sum
```

The total number of observations in the resulting dataset is 524, key variables include the dumpster number, the date, weight in tons, volume in cubic yards, type of the trash, and homes powered. The total weight of trash collected by Professor Trash Wheel is 135.5 tons. 
# What was the total number of sports balls collected by Mr. Trash Wheel in 2020?

## Problem 3
